---
title: "Module 7: Advanced SEM"
subtitle: "Invariance"
output: html_notebook
---

```{r}
library(tidyverse)
library(lavaan)
```


Items from the CES-D Somatic and Retarded Activity factor

- cesdOl. I was bothered by things that usually don't bother me. 
- cesdO2. I did not feel like eating; my appetite was poor.
- cesdO7. I felt that everything I did was an effort
- cesdll. My sleep was restless
- cesd20. I could not get "going."

Our results will not be identical b/c we cannot use robust estimators---we need the original data to do that. So can only compare to the ML estimates, which we replicate.

```{r}

var_names <- c("cesd01", "cesd02", "cesd07", "cesd11", "cesd20")

cov_black <- lav_matrix_lower2full(
  c(
  0.49587,
  0.16935, 0.57562,
  0.19368, 0.17265, 1.29030,
  0.14082, 0.13106, 0.31683, 0.87469,
  0.18769, 0.12476, 0.34361, 0.26508, 0.66199
  )
)

rownames(cov_black) <- colnames(cov_black) <- var_names

means_black <- c(0.35081, 0.34677, 0.80645, 0.58871, 0.45565)

names(means_black) <- var_names

cov_white <-  lav_matrix_lower2full(
  c(
  0.49822,
  0.12634, 0.42714,
  0.20046, 0.17796, 0.76513,
  0.18179, 0.17049, 0.27322, 0.82264,
  0.16644, 0.16163, 0.31836, 0.24146, 0.63059
  )
)

rownames(cov_white) <- colnames(cov_white) <- var_names

means_white <- c(0.32834, 0.24002, 0.48553, 0.61228, 0.43713)
names(means_white) <- var_names
```

## Dimensional/configural model

```{r}
combined_cov <- list(black = cov_black, white = cov_white)
combined_means <- list(black = means_black, white = means_white)

m_config <- '
somatic =~ cesd01 + cesd02 + cesd07 + cesd11 + cesd20
'

f_config <- cfa(m_config, sample.cov=combined_cov, sample.mean = combined_means, meanstructure=TRUE, sample.nobs = c(248, 2004))

summary(f_config)
```

## Metric model

```{r}
m_metric <- '
somatic =~ c(1,1)*cesd01 + c(b2,w2)*cesd02 + c(b3,w3)*cesd07 + c(b4,w4)*cesd11 + c(b5,w5)*cesd20
'

f_metric <- cfa(
  m_metric, 
  sample.cov=combined_cov, 
  sample.mean = combined_means, 
  meanstructure=TRUE, 
  sample.nobs = c(248, 2004)
  )

summary(f_metric)
```

```{r}
m_strong <- '
somatic =~ cesd01 + cesd02 + cesd07 + cesd11 + cesd20
'

f_strong <- cfa(
  m_strong, 
  sample.cov=combined_cov, 
  sample.mean = combined_means, 
  meanstructure=TRUE, 
  sample.nobs = c(248, 2004),
  group.equal = c("loadings", "intercepts")
  )

summary(f_strong)
```

```{r}
AIC(f_config, f_metric, f_strong)
```

```{r}
residuals(f_strong, type="cor")
```

```{r}
# sometimes easier to get it to pop out
lavTestScore(f_strong)
```

## Partial strong invariance

```{r}
m_p_strong <- '
somatic =~ cesd01 + cesd02 + cesd07 + cesd11 + cesd20
cesd07 ~ c(ta,tb)*1
'

f_p_strong <- cfa(
  m_p_strong, 
  sample.cov=combined_cov, 
  sample.mean = combined_means, 
  meanstructure=TRUE, 
  sample.nobs = c(248, 2004),
  group.equal = c("loadings", "intercepts")
  )

summary(f_p_strong)
```

```{r}
AIC(f_config, f_metric, f_strong, f_p_strong)
```

## partial strict invariance

```{r}
m_p_strict <- '
somatic =~ cesd01 + cesd02 + cesd07 + cesd11 + cesd20
cesd07 ~ c(ta,tb)*1
cesd07 ~~ c(ra, rb)*cesd07
cesd02 ~~ c(rc, rd)*cesd02
'

f_p_strict <- cfa(
  m_p_strict, 
  sample.cov=combined_cov, 
  sample.mean = combined_means, 
  meanstructure=TRUE, 
  sample.nobs = c(248, 2004),
  group.equal = c("loadings", "intercepts", "residuals")
  )

summary(f_p_strict)
```

## Equal factor means

```{r}
m_efm <- '
somatic =~ cesd01 + cesd02 + cesd07 + cesd11 + cesd20
cesd07 ~ c(ta,tb)*1
cesd07 ~~ c(ra, rb)*cesd07
cesd02 ~~ c(rc, rd)*cesd02
'

f_efm <- cfa(
  m_efm, 
  sample.cov=combined_cov, 
  sample.mean = combined_means, 
  meanstructure=TRUE, 
  sample.nobs = c(248, 2004),
  group.equal = c("loadings", "intercepts", "residuals", "means")
  )

summary(f_efm)
```

```{r}
m_efv <- '
somatic =~ cesd01 + cesd02 + cesd07 + cesd11 + cesd20
cesd07 ~ c(ta,tb)*1

cesd07 ~~ c(ra, rb)*cesd07
cesd02 ~~ c(rc, rd)*cesd02
'

f_efv <- cfa(
  m_efv, 
  sample.cov=combined_cov, 
  sample.mean = combined_means, 
  meanstructure=TRUE, 
  sample.nobs = c(248, 2004),
  group.equal = c("loadings", "intercepts", "residuals", "means", "lv.variances")
  )

summary(f_efv)
```

```{r}
AIC(
  f_config,
  f_metric,
  f_strong,
  f_p_strong,
  f_p_strict,
  f_efm,
  f_efv
)
```
