---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

## Variable assignment

Make a new variable called `x` equal to the number 3

```{r}

```


Make a second variable called `y` equal to the number 5

```{r}

```

Make a third variable `z` that is the product of `x` and `y`

```{r}

```

Calculate the sum of `x`, `y`, and `z` but do not assign it to a variable, just let it be printed by the chunk below.

```{r}

```

Change the variable `x` so that it is equal to 9.

```{r}

```

Make a second variable called `z2` that is the product of  `x` and `y`. Notice how this compares to `z`.

```{r}

```

Run the code below, then at the console, enter `y`. What happened? Ask yourself why.

```{r}
y * 10
```


## Working with data (tibbles)

Read in the file `example-data.csv` as a tibble or dataframe called `df`.

```{r}

```

View the data you have read in in whichever way you prefer (e.g. `view()`, `glimpse()`, `head()`, or even `print()`).

Run the code below and then look at `df` again. Has anything changed? Ask yourself why or why not.

```{r}
# df %>% 
#   mutate(x, y, z)
```

Rewrite the code chunk above so that the changes actually affect `df`.

```{r}

```

Uncomment the code in the chunk below and run it. What happens? 

```{r}

```

Ask yourself: What's going on here? 

Create a new variable `xyz` that contains the `xyz` values in the `df` data.

```{r}

```

