---
title: "Introduction to R"
output: html_notebook
---

# loading packages

```{r}
# loading packages with library() <-- also note # for comments
library(tidyverse) # ggplot2, readr, dplyr, etc.
```

# loading data

```{r}
# loading data
df <- read_csv("stroop-2014.csv")
```

# inspect data + indexing dataframes / grabbing values

```{r}
head(df)
```

e.g. check column types, find missing values, find the largest value, smallest value, etc

```{r}
any(is.na(df))
```

```{r}
mean(df$congruent)
min(df$congruent)
max(df$congruent)
```

```{r}
mean(df$incongruent)
min(df$incongruent)
max(df$incongruent)
```

# define some new functions

* `se()`
* `standardize()`

```{r}
se <- function(x){sd(x)/sqrt(length(x))}

standardize <- function(x){
  (x - mean(x, ...))/sd(x, ...)
}
```

```{r}
se <- function(x){sd(x)/sqrt(length(x))}


sd(df$congruent)

se(df$congruent)
```

# Data munging and visualization

The "pipe" operator ` %>% ` is handy.

```{r}
# to long format
df_long <- df %>% 
  gather(trial_type, rt, congruent, incongruent)

# see what happened
df_long %>% 
  arrange(subj_num) %>% 
  head()
```

```{r}
# making a boxplot
df_long %>% 
  ggplot(aes(x=trial_type, y=rt)) +
  geom_boxplot()
```

```{r}
# summarizing
df_long %>% 
  group_by(trial_type) %>% 
  summarize(mean_rt = mean(rt), se_rt = se(rt))
```

```{r}
# plotting means and standard errors
df_long %>% 
  group_by(trial_type) %>% 
  summarize(mean_rt = mean(rt), se_rt = se(rt)) %>% 
  mutate(ul = mean_rt + se_rt, ll = mean_rt- se_rt) %>% 
  ggplot(aes(x=trial_type, y = mean_rt, ymin=ll, ymax=ul)) +
  geom_pointrange()
```

# do a statistical test

```{r}
t.test(df$congruent, df$incongruent, paired=TRUE)
```

`t.test()` returns a "named list." This can be indexed by its names with `$` or with double brackets `[[ ]]`.

```{r}
result <- t.test(df$congruent, df$incongruent, paired=TRUE)

result$p.value
result[["statistic"]]
```


# Modify and transform data

```{r}
# filter out a "bad" subject

df %>% 
  filter(subj_num != 3)
```

```{r}
# translate seconds to ms -- but note we don't have accuracy, so seconds is probably better
df %>% 
  mutate(
    congruent_ms = congruent*1000,
    incongruent_ms = incongruent*1000
  )
```

```{r}
# renaming: capitalize condition names
df %>% 
  rename(Incongruent = incongruent, Congruent = congruent)

```

```{r}
# editing variables, 3 different ways

df_long %>% 
  mutate(trial_type = ifelse(trial_type=="congruent", "Congrent", "Incongruent"))

df_long %>% 
  mutate(
    trial_type = fct_recode(
      trial_type, 
      "Congruent"="congruent", 
      "Incongruent"="incongruent")
    )

df_long %>% 
  mutate(
    trial_type = case_when(
      trial_type=="congruent" ~ "Congruent",
      trial_type=="incongruent" ~ "Incongruent"
    )
  )
```

```{r}
# creating a new categorical variable

df %>% 
  mutate(avg = (congruent + incongruent)/2 ) %>% 
  mutate(
    speed_group = case_when(
      percent_rank(-avg) > .67 ~ "fast",
      percent_rank(-avg) > .33 ~ "medium",
      TRUE ~ "slow"
    )
  ) %>% 
  gather(trial_type, rt, contains("congruent")) %>% 
  ggplot(aes(x=speed_group, y = rt, fill=trial_type)) +
  geom_boxplot()
```

## histograms

```{r}
df_long %>% 
  ggplot(aes(x=rt, fill=trial_type)) +
  geom_histogram(alpha=.7, position="identity")
```

## Publication quality figures

```{r}
# make a pretty pointrange plot / also make a bar plot
```


```{r}
corr_val <- round(cor(df$congruent, df$incongruent), 3)

df %>% 
ggplot(aes(x = congruent, y = incongruent)) + 
  geom_smooth(method="lm", alpha=.2) +
  geom_point() +
  annotate("text", x = 20, y = 13, label=paste("r =", corr_val), size = 4) +
  theme_bw(base_size=12) +
  theme(
    aspect.ratio=1,
    panel.grid = element_blank()
    ) +
  labs(
    x = "Congruent RT (s)", 
    y = "Incongruent RT (s)", 
    title = "Scatterplot of Stroop task response times"
    )
```

# A larger example

this data has 21k rows and factors have many levels. But notice how our code doesn't really change!

```{r}
# I should get some of my own GSS data with more interesting variables
data(gss_cat)

glimpse(gss_cat)
```

```{r}
se <- function(x,...){sd(x,...)/sqrt(length(x))}

gss_cat %>% 
  group_by(marital) %>% 
  summarize(M = mean(tvhours, na.rm=TRUE), SE = se(tvhours, na.rm=TRUE)) %>% 
  ggplot(aes(x=marital, y = M, ymin = M - SE, ymax=M+SE)) +
  facet_wrap(~marital) +
  geom_pointrange()
```
```{r}
gss_cat %>% 
  ggplot(aes(x=age, y = tvhours)) +
  # facet_wrap(~marital) +
  geom_smooth() +
  geom_jitter(alpha=.1)
```
[do a `lm()` on something of some interest]


# A real example (DeJesus children eating food example)

```{r}
# load dejesus data
```

Make this an in-class exercise

Illustrates need to inspect data, mutate based on another variable. Check for impossible values and fix them. Then deal with NAs.

Want them to:
* clean up variable naming
* deal with NAs
* notice impossible values
* fix them carefully (check if what we do is safe)
* finally, make a plot
* maybe, do a test

